---
title: "Performance Testing and Comparative Benchmarking for Data.Table"
author: "Doris Amoakohene"
date: "`r Sys.Date()`"
output:
  powerpoint_presentation: default
  beamer_presentation: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```


## Outline of Presentation 

-Introduction
-Understanding Performance Regressions in data.table
-Introduction to the atime package for performance analysis
-Comparing data.table with Python Packages
-Enhancing data.table: Educational Resources
-Conclusion

## Why Benchmarking?

- data.table is an extension of R’s data.frame, designed to handle large datasets efficiently

- Several major version changes have been introduced, aiming to enhance speed and efficiency. These changes include algorithmic optimizations, memory management improvements, and enhancements to parallel processing capabilities.

- To evaluate data.table performance, it is essential to employ benchmarking methodologies. The approach, I used to utilize the atime and  atime_versions function from the atime package, which measures the actual execution time of specific operations.

## Why Benchmarking?

- The time function allows to compare and visualize the asymptotic performance (time and memory usage) of the different functions 

- The atime_versions allows for accurate comparisons between different versions of the data.table package, by benchmarking against time and memory usage and also giving a graphical visualization of the results

- In most benchmarking techniques, such as airspeed velocity, Conbench, and Touchstone, a single or constant value of N is typically used to measure performance. However, when it comes to benchmarking the performance of data.table, N can vary or follow a specific sequence depending on the number of data points being analyzed. This approach allows researchers and developers to evaluate the scalability and efficiency of data.table under different data sizes, providing a more comprehensive understanding of its performance characteristics. 


```{r}
atime::atime_versions( 
  pkg.path = tdir, 
  pkg.edit.fun = pkg.edit.fun, 
  N = 10^seq(1,20), 
  setup = { 
    set.seed(108) 
    d <- data.table( 
      id3 = sample(c(seq.int(N*0.9), 
                     sample(N*0.9, N*0.1, TRUE))), 
      v1 = sample(5L, N, TRUE), 
      v2 = sample(5L, N, TRUE)) 
  }, 
  expr = data.table:::`[.data.table`(d, , (max(v1)-min(v2)), by = id3),
  "Before" = "15f0598b9828d3af2eb8ddc9b38e0356f42afe4f",
 "Regression" = "6f360be0b2a6cf425f6df751ca9a99ec5d35ed93",
 "Fixed" = "ba32f3cba38ec270587e395f6e6c26a80be36be6"
```


